---
title: "QSURE Statistical analysis"
author: "Michael and Karissa"
knit: (function(inputFile, encoding) { 
      out_dir <- 'output';
      rmarkdown::render(inputFile,
                        encoding=encoding, 
                        output_file=file.path(here::here(), out_dir, 'analysis.html')) })
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.align = 'center',
	fig.width = 12,
	fig.height = 10
)

library(tidyverse)
library(gtsummary)
library(survival)
library(survminer)
library(here)


load(file = here(
  "data", "diabetes.RData"))


```


## Summarizing data 

This is a common table 1 you would often see in manuscripts. Below I am staratifying the data by gender and adding a p value. The specific tests will be specified in the footnote. You can remove the p-values, and overall column if you do not want them there. 

```{r}

tbl_summary(df
            ,include = -id
            ,by = gender) %>% 
  bold_labels() %>% 
  add_overall() %>% 
  add_p()
      

```

### Summarize continious data

Another common way to summarize data would be to summarize a continuous variable by one or more categorical variables. 

```{r}
tbl_continuous(df,
               variable = stabilized_glucose,
               by = gender,
               include = location)


```


### Reporting data with ggplot2

`ggplot2` is one of the most common ways to visualize data in R. 
<br>
Components of ggplot:
<br>
- data: data.frame/tibble 
- geom_* :type of plot (* means there are multiple options [e.g - geom_bar, geom_point])
  - these geometric objects tell ggplot what kind of graph you are building 
  - scatter plot, box plot, bar plot, etc...
- aesthetics (aes): tells ggplot how to represent the data
- stats: can transform data to be plotted 
- theme: handles general graphing options (color background, font sizes)
- scale: controls options between data and aes, plot limits, colors, line types, etc...
<br>
Necessary components of ggplot
<br>
- only need data, geom_* and aesthetics to make a graph, the rest are augments of this base plot 

```{r}

sumdf <- df %>% 
          group_by(location, gender) %>% 
          summarise(age = mean(age, na.rm = TRUE)) %>% 
          ungroup() %>% 
          mutate(var = paste(location, gender, sep  = " ")) %>% 
          filter(!is.na(location))


ggplot(sumdf) +
    geom_bar(aes(x= var, y = age, fill = var),stat = "identity") +
    scale_x_discrete(" ") +
    scale_fill_manual(" ", values = c("black","blue","grey","red"))+
    scale_y_continuous("Age", limits = c(0,60)) +
    theme_bw()

```

You can also represent the data in a more sophisticated way with facets. 


```{r}

ggplot(sumdf) +
    geom_bar(aes(x= location, y = age, fill = location),stat = "identity") +
    facet_grid(.~gender)+
    scale_x_discrete(" ") +
    scale_fill_manual(" ", values = c("black","blue"))+
    scale_y_continuous("Age", limits = c(0,60)) +
    theme_bw()

```


There are many types of geom_* and it is best to look at them from the https://ggplot2.tidyverse.org/ website. 